New lower bounds for Schur and weak Schur numbers

### 2021-04-06 article/article.tex
avancement sat

\begin{definition}
A subset \(A\) of \(\mathbb{N}\) is said to be \textit{sum-free} when :
\[ \forall (a,b) \in A^2 \text{, } a+b \notin A\]
\end{definition}

@@

Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$. If there exists a partition of $r$ weakly sum-free subsets of $[\![1,n]\!]$ and a partition of $s$ sum-free subsets of $[\![1,p]\!]$ then there exists a partition of $r+s$ weakly sum-free subsets of $[\![1,p(n+\left \lceil \frac{n}{2} \right \rceil + 1)+n]\!]$

@@

In the above formula, every color plays a symmetric role. Hence the search space can reduced by \(k!\) by ordering the subsets, that is by 
enforcing that \(m_1 < ... < m_k\). The correspond clauses are: \linebreak
\textbf{symmetry breaking:}  \(x^{(1)}_1 = 1\) and \(\forall c \in [\![2,k-1]\!], \forall i \in [\![1,WS(c - 1)+1]\!], x^{(1)}_c \lor ... \lor x^{(i)}_c \lor \neg x^{(i+1)}_{c+1}\)

\subsection{The search space previously used in computer search for lower bound may not contain the optimal partitions}

Rowley's new lower bound for \(WS(6)\) (642)  \hyperlink{label2}{\textbf{[2]}} was a quite significant improvement upon the former 
best known lower bound (582). This previous lower bound was found using a computer with Monte-Carlo methods and by making the 
assumption that a good partition for \(WS(n+1)\) starts with a good partition for \(WS(n)\) which is true for small values of \(n\).
Therefore, one may wonder whether the limiting factor is the assumption or the methods used to search for partitions. It appears that 
the latter is correct.

@@

\par
Every weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets starts with the following sequence 1121222133. Then 11 is always 
either in subset 1 or 3, 12 is always in subset 3 and so. For every integer in \([\![1,65]\!]\), we computed in which subset it can appear. 
By using this constraints, we could then compute for every integer in \([\![1,185]\!]\), in which subset it can appear in a weakly sum-free 
partition of  \([\![1,185]\!]\) which starts with a  weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets. Adding these constraints to the 
formula corresponding to the above theorem gives additional information to the SAT solver without losing in generality.

\subsection{Weak Schur number five}

>>>>

\begin{definition}
A subset \(A\) of \(\mathbb{N}\) is said to be \textit{sum-free} when:
\[ \forall (a,b) \in A^2 \text{, } a+b \notin A\]
\end{definition}

@@

Let $(n,p), (r,s) \in (\mathbb{N}^*)^2$. If there exists a partition of $r$ weakly sum-free subsets of $[\![1,n]\!]$ and a partition of $s$ sum-free 
subsets of $[\![1,p]\!]$ then there exists a partition of $r+s$ weakly sum-free subsets of $[\![1,p(n+\left \lceil \frac{n}{2} \right \rceil + 1)+n]\!]$

@@

In the above formula, every color plays a symmetric role. Hence the search space can reduced by \(k!\) by ordering the subsets, that is by 
enforcing that \(m_1 < ... < m_k\). The corresponding clauses are: \linebreak
\textbf{symmetry breaking:}  \(x^{(1)}_1 = 1\) and \(\forall c \in [\![2,k-1]\!], \forall i \in [\![1,WS(c - 1)+1]\!], x^{(1)}_c \lor ... \lor x^{(i)}_c \lor \neg x^{(i+1)}_{c+1}\)

\subsection{The search space previously used in computer search for lower bounds may not contain the optimal partitions}

Rowley's new lower bound for \(WS(6)\) (642)  \hyperlink{label2}{\textbf{[2]}} was a quite significant improvement upon the former 
best known lower bound (582). This previous lower bound was found using a computer (often with Monte-Carlo) methods and by making the 
assumption that a good partition for \(WS(n+1)\) starts with a good partition for \(WS(n)\) which is true for small values of \(n\).
Therefore, one may wonder whether the limiting factor is the assumption or the methods used to search for partitions. It appears that 
the latter is correct.

@@

\par
Every weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets starts with the following sequence 1121222133. Then 11 is always 
either in subset 1 or 3, 12 is always in subset 3 and so on. For every integer in \([\![1,65]\!]\), we computed in which subset it can appear. 
By using this constraints, we could then compute for every integer in \([\![1,185]\!]\), in which subset it can appear in a weakly sum-free 
partition of  \([\![1,185]\!]\) which starts with a  weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets. Adding these constraints to the 
formula corresponding to the above theorem gives additional information to the SAT solver without losing in generality.

\par
The above theorem shows that the previous lower bound for \(WS(6)\) is optimal in the search space considered by the papers which found it. 
Therefore, finding a partition of \([\![1,n]\!]\) in 6 weakly sum-free subsets for some \(n \geq 590\) which does not have a template-like structure 
would be extremely interesting since it could give indications on a new search space for improving lower bounds with a computer. More generally, 
it questions the search space previously used for finding lower bounds for \(WS(n)\) with a computer. In particular, to our knowledge every paper 
that found the lower bound \(WS(5) \geq 196\) used this assumption [reference necessaire]. Therefore one may wonder if this actually a good lower 
bound. In the next subsection, we give properties that a partition of \([\![1,197]\!]\) in 5 weakly sum-free subsets has to verify.

\subsection{Weak Schur number five}
As explained in the previous subsection, the search space used for showing that  \(WS(5) \geq 196\) may not contain optimal solution. In this subsection, 
we give necessary conditions for a hypothetical partition of \([\![1,197]\!]\) in 5 weakly sum-free subsets using the same method as the previous subsection.
